---
title: "Salmon group 8"
author: "Nicole Schmidt"
date: "2/2/2021"
output: html_document
---

### Overview

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Attach packages and read in and tidy data 

library(tidyverse)
library(janitor)
library(here)
library(lubridate)

willamette_fish_passage <- read_csv(here("data", "willamette_fish_passage.csv")) %>% 
  clean_names() %>% 
  select(coho, jack_coho, steelhead, date) %>% 
  mutate(date = lubridate::mdy(date))

# converting to long format to put all fish species in one column
willamette_long <- willamette_fish_passage %>% 
  pivot_longer(1:3,
               names_to = "fish_species",
               names_prefix = NULL,
               values_drop_na = FALSE) # decided not to drop observations with n/a values becuase in Tab 1 it says to change na values to zero
  # replace na values with zero
willamette_long[is.na(willamette_long)] = 0 

```

### Tab 1: Original time series - (Nicole)

### Tab 2: Seasonplots - (Mukta)

### Tab 3: Annual counts by species - Grace
This “Summary statistics and analysis” tab should contain: 
A finalized figure of annual totals for fish passage, for each of the 3 species (coho, jack coho, steelhead). You decide if this is in a single panel, or multiple panels. Add a figure caption. 
Add 2 - 3 bulletpoints summarizing major trends you see in the annual totals by species from 2000 - 2010.
Remember: All code that you used to wrangle the data and prepare the graphs should be available to see if we click on the Code buttons. 
```{r}


# summarize count by species and year
fish_year <- willamette_long %>% 
  mutate(year = year(date))  %>% # pulling just the year from the date
  replace_na() %>% 
  group_by(year, fish_species) %>% 
  summarize(annual_total = sum(value))
  
# graphing fish passage by species and year
ggplot(data = fish_year,
       aes(x = year,
           y = annual_total,
           color = fish_species)) +
  geom_line() +
  scale_x_continuous(expand = c(0,0),
                     breaks = seq(2001,2010,1)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "Year",
       y = "Annual observed",
       title = "Annual fish passage totals by species") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 0.5 ))
  #theme(legend.title = element_text(color = "blue"))
  #theme(axis.title.x = element_text(hjust = 0.5))
  #scale_fill_manual(values = c("#21908c80", "#fde72580", "yellow"), # custom colors
  #                  name = "Fish pecies",
  #                   labels = c("Coho", "Jack Coho", "Steelhead"))

```





