---
title: "Salmon group 8"
author: "Nicole Schmidt"
date: "2/2/2021"
output: html_document
---

### Overview

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Attach packages and read in and tidy data 

library(tidyverse)
library(janitor)
library(here)
library(lubridate)

willamette_fish_passage <- read_csv(here("data", "willamette_fish_passage.csv")) %>% 
  clean_names() %>% 
  select(coho, jack_coho, steelhead, date) %>% 
  mutate(date = lubridate::mdy(date))

# converting to long format to put all fish species in one column
willamette_long <- willamette_fish_passage %>% 
  pivot_longer(1:3,
               names_to = "fish_species",
               names_prefix = NULL,
               values_drop_na = FALSE) # decided not to drop observations with n/a values becuase in Tab 1 it says to change na values to zero
  

```

### Tab 1: Original time series - (Nicole or Mukta?)

### Tab 2: Seasonplots - (Nicole or Mukta?)

### Tab 3: Annual counts by species - Grace
This “Summary statistics and analysis” tab should contain: 
A finalized figure of annual totals for fish passage, for each of the 3 species (coho, jack coho, steelhead). You decide if this is in a single panel, or multiple panels. Add a figure caption. 
Add 2 - 3 bulletpoints summarizing major trends you see in the annual totals by species from 2000 - 2010.
Remember: All code that you used to wrangle the data and prepare the graphs should be available to see if we click on the Code buttons. 
```{r}



```





